---
title: "WUP_04"
author: "Shael Quintero"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file = "GSScsv.csv")

```
```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
head(mydata)
```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Comparing the GSS data set's numerical and categorical data using the ChatGPT and CandP auxiliary files. We will contrast WORDSUM and examine the impact of educational level.


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r} 
ggplot(data = mydata, aes(x = DEGREE, y = WORDSUM)) +
  geom_boxplot(notch = FALSE, outlier.colour = "red", outlier.size = 2, fill = "darkgray") +
  labs(x = "DEGREE", y = "WORDSUM SCORE", title = "WORDSUM EXPLAINED BY DEGREE") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```
The IQR of the high school and LT High School appears to be lower overall than the median of all college groupings.



### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats( WORDSUM ~ DEGREE , data = mydata)
```
The LT Highschool and high school subcategories have mean and median values that are lower than those of the data points collected at the university level.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov( WORDSUM ~ DEGREE , data = mydata)
summary(model1)
```
```{r}
par(mar = c(3,15,2,2))
plot(TukeyHSD(model1,conf.level = 0.99),las=2)             
```                                                                   


# Discussion and Conclusion

To compare mean values between different groups in an ANOVA, the Tukey HSD test is frequently employed. It offers a mechanism of determining which particular pairs of groups have significantly differing means.The visually pleasing plot and complete visibility of the labels are both made possible by the generous left margin established by par(mar = c(3,15,2,2)).
The intervals shown in the plot can be trusted to be accurate because a 99% confidence level was chosen.


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
